---
title: "Logistic Regression Classification - Full"
author: "Mervin M Fansler"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_float: true
---

# Purpose

Here we fit all data with the logistic regression model, without any test-train
split since we are interested in all the information in the data.

# Initialization

## Libraries
```{r libs, message=FALSE, warning=FALSE}
library(magrittr)
library(tidyverse)
library(ggtern)
library(broom)
library(caret)
library(nnet)
library(readxl)
library(writexl)
```

## Parameters
```{r set_params}
set.seed(20220607)
N_STEPS=500

RDS_IN="data/df_tiger_clean.Rds"

LOC_COLORS=c("ER"="#4B74A6", "TG"="#E4E4CA", "CY"="#949494", "DF"="#CCCCCC")
LOC_COLORS_DARK=c("ER"="#4B74A6", "TG"="#C9C497", "CY"="#949494", "DF"="#CCCCCC")

COVARIATES=c(
    "Leb_HuR_CLIP", 
    "Muk_HuR_CLIP", 
    "CPE_annoUTR", 
    "CCCCCC_3UTR_num",
    "Pumilio_3UTR_num", 
    "PUM2_CLIP", 
    "KHSRP_CLIP", 
    "CPSF6_CLIP", 
    "TIS11B_CLIP", 
    "TIA1_L1_CLIP", 
    "PTBP1_CLIP", 
    "HNRNPC_CLIP",
    "CELF2_CLIP", 
    "HNRNPM_CLIP", 
    "HNRNPD_CLIP", 
    "Zhang_RBM15_RIP",
    "LIN28B_CLIP", 
    "IGF2BP3_CLIP", 
    "TARDBP_CLIP", 
    "TAF15_CLIP",
    "FUS_CLIP", 
    "METAP2_CLIP", 
    "DDX6_CLIP", 
    "EIF4A3_CLIP",
    "LARP4B_exon_CLIP", 
    "hnRNPK_CLIP", 
    "GEMIN5_CLIP", 
    "CSTF2T_CLIP",
    "DDX24_CLIP", 
    "PCBP2_CLIP", 
    "PUM_RIP12",
    "Anno_3UTR_length")
```

# Data
## Loading
```{r load_data, message=FALSE}
df_tiger <- readRDS(RDS_IN)
```

## Preprocessing
### Subset
```{r prepare_data}
df_data <- select(df_tiger, category, all_of(COVARIATES))
```

### Imputing
```{r data_impute}
df_imputed <- df_data %>%
    mutate_at("Anno_3UTR_length", ~ ifelse(is.na(.x), median(.x, na.rm=TRUE), .x)) %>%
    mutate_at(-1, ~ ifelse(is.na(.x), 0, .x))
```

### Data Normalizing
```{r data_normalize}
df_scaled <- mutate_at(df_imputed, -1, ~ scale(sqrt(.x))[,1])
```

### Counts
```{r category_cts}
count(df_scaled, category) %>%
    knitr::kable()
```

# Analysis
## Logistic Regression Model
### Fit
```{r fit_model}
model <- multinom(category ~ ., data=df_scaled, maxit=1000)
AIC(model)
tidy(model) %>% 
    arrange(p.value) %>%
    knitr::kable()
```

### Export
```{r export_fit} 
outfile <- sprintf("tbl/%s-model-fit-lr-all.xlsx", format(Sys.time(), '%Y%m%d'))
tidy(model) %>%
    arrange(p.value) %>%
    write_xlsx(outfile)
```


### Model Accuracy
```{r prediction_accuracy}
mean(predict(model, df_scaled) == df_scaled$category)
```

## Confusion Matrix
```{r plot_predicted_confusion, fig.width=5, fig.height=4}
df_scaled %>%
    mutate(predicted=predict(model, .)) %>%
    count(category, predicted) %>% 
    ggplot(aes(category, predicted, fill=n)) +
    geom_tile() + geom_text(aes(label=n)) +
    scale_fill_gradient(low="white") +
    labs(title="All Data", x="True Category", y="Predicted Category") +
    theme_light()

df_scaled %>%
    mutate(predicted=predict(model, .)) %$%
    confusionMatrix(predicted, category)
```

### Coefficients
```{r plot_coefficients, fig.width=6, fig.height=8}
coefficients(model) %>%
    as_tibble(rownames="category") %>%
    select(-c("(Intercept)")) %>%
    pivot_longer(-c("category"), names_to="variable", values_to="value") %>%
    mutate(variable=fct_reorder(variable, abs(value), max)) %>%
    ggplot(aes(y=variable, x=value, fill=category)) +
    geom_vline(xintercept=0) +
    geom_bar(stat='identity', position='dodge', color='black', size=0.1, width=0.9) +
    scale_fill_manual(values=LOC_COLORS) +
    labs(x="Coefficient", y=NULL, fill="Category") +
    theme_light()

ggsave("img/lr-all-coefs-all.pdf", width=6, height=8, dpi=300)
```

### Coefficient Pairs
#### TIAL1, TIS11B
```{r coefs_tia_tis, fig.width=5, fig.height=4}
df_scaled %>% 
    summarize(across(everything(), mean)) %>% 
    slice(rep(1, N_STEPS)) %>%
    mutate(TIA1_L1_CLIP=seq(-10, 10, length.out=N_STEPS)) %>%
    slice(rep(1:N_STEPS, times=N_STEPS)) %>%
    mutate(TIS11B_CLIP=rep(seq(-10,10, length.out=N_STEPS), each=N_STEPS)) %>%
    mutate(category=predict(object=model, .)) %>%
    ggplot(aes(x=TIA1_L1_CLIP, y=TIS11B_CLIP, fill=category)) +
    geom_tile(linejoin="bevel") +
    scale_fill_manual(values=LOC_COLORS) +
    theme_bw()

ggsave("img/lr-all-pairs-tial1-tis11b.pdf", width=5, height=4, dpi=300)
```

#### Pumilio Num, TIS11B
```{r coefs_pum_tis, fig.width=5, fig.height=4}
df_scaled %>% 
    summarize(across(everything(), mean)) %>% 
    slice(rep(1, N_STEPS)) %>%
    mutate(Pumilio_3UTR_num=seq(-10, 10, length.out=N_STEPS)) %>%
    slice(rep(1:N_STEPS, times=N_STEPS)) %>%
    mutate(TIS11B_CLIP=rep(seq(-10,10, length.out=N_STEPS), each=N_STEPS)) %>%
    mutate(category=predict(object=model, .)) %>%
    ggplot(aes(x=Pumilio_3UTR_num, y=TIS11B_CLIP, fill=category)) +
    geom_tile(linejoin="bevel") +
    scale_fill_manual(values=LOC_COLORS) +
    theme_bw()

ggsave("img/lr-all-pairs-pumnum-tis11b.pdf", width=5, height=4, dpi=300)
```

#### Pum1/2 RIP, TIS11B
```{r coefs_pum12_tis, fig.width=5, fig.height=4}
df_scaled %>% 
    summarize(across(everything(), mean)) %>% 
    slice(rep(1, N_STEPS)) %>%
    mutate(PUM_RIP12=seq(-10, 10, length.out=N_STEPS)) %>%
    slice(rep(1:N_STEPS, times=N_STEPS)) %>%
    mutate(TIS11B_CLIP=rep(seq(-10,10, length.out=N_STEPS), each=N_STEPS)) %>%
    mutate(category=predict(object=model, .)) %>%
    ggplot(aes(x=PUM_RIP12, y=TIS11B_CLIP, fill=category)) +
    geom_tile(linejoin="bevel") +
    scale_fill_manual(values=LOC_COLORS) +
    theme_bw()

ggsave("img/lr-all-pairs-pum12rip-tis11b.pdf", width=5, height=4, dpi=300)
```

#### Pum1/2 RIP, TIAL1
```{r coefs_pum12_tial1, fig.width=5, fig.height=4}
df_scaled %>% 
    summarize(across(everything(), mean)) %>% 
    slice(rep(1, N_STEPS)) %>%
    mutate(PUM_RIP12=seq(-10, 10, length.out=N_STEPS)) %>%
    slice(rep(1:N_STEPS, times=N_STEPS)) %>%
    mutate(TIA1_L1_CLIP=rep(seq(-10,10, length.out=N_STEPS), each=N_STEPS)) %>%
    mutate(category=predict(object=model, .)) %>%
    ggplot(aes(x=PUM_RIP12, y=TIA1_L1_CLIP, fill=category)) +
    geom_tile(linejoin="bevel") +
    scale_fill_manual(values=LOC_COLORS) +
    theme_bw()

ggsave("img/lr-all-pairs-pum12rip-tial1.pdf", width=5, height=4, dpi=300)
```

#### Pumilio Num, TIAL1
```{r coefs_pum_tia, fig.width=5, fig.height=4}
df_scaled %>% 
    summarize(across(everything(), mean)) %>% 
    slice(rep(1, N_STEPS)) %>%
    mutate(Pumilio_3UTR_num=seq(-10, 10, length.out=N_STEPS)) %>%
    slice(rep(1:N_STEPS, times=N_STEPS)) %>%
    mutate(TIA1_L1_CLIP=rep(seq(-10,10, length.out=N_STEPS), each=N_STEPS)) %>%
    mutate(category=predict(object=model, .)) %>%
    ggplot(aes(x=Pumilio_3UTR_num, y=TIA1_L1_CLIP, fill=category)) +
    geom_tile(linejoin="bevel") +
    scale_fill_manual(values=LOC_COLORS) +
    theme_bw()

ggsave("img/lr-all-pairs-pumnum-tial1.pdf", width=5, height=4, dpi=300)
```

#### TIS11B, LARP4B
```{r coefs_tis_larp, fig.width=5, fig.height=4}
df_scaled %>% 
    summarize(across(everything(), mean)) %>% 
    slice(rep(1, N_STEPS)) %>%
    mutate(TIS11B_CLIP=seq(-10, 10, length.out=N_STEPS)) %>%
    slice(rep(1:N_STEPS, times=N_STEPS)) %>%
    mutate(LARP4B_exon_CLIP=rep(seq(-10,10, length.out=N_STEPS), each=N_STEPS)) %>%
    mutate(category=predict(object=model, .)) %>%
    ggplot(aes(x=TIS11B_CLIP, y=LARP4B_exon_CLIP, fill=category)) +
    geom_tile(linejoin="bevel") +
    scale_fill_manual(values=LOC_COLORS) +
    theme_bw()

ggsave("img/lr-all-pairs-tis11b-larp4b.pdf", width=5, height=4, dpi=300)
```

#### TIA1L1, LARP4B
```{r coefs_tia_larp, fig.width=5, fig.height=4}
df_scaled %>% 
    summarize(across(everything(), mean)) %>% 
    slice(rep(1, N_STEPS)) %>%
    mutate(TIA1_L1_CLIP=seq(-10, 10, length.out=N_STEPS)) %>%
    slice(rep(1:N_STEPS, times=N_STEPS)) %>%
    mutate(LARP4B_exon_CLIP=rep(seq(-10,10, length.out=N_STEPS), each=N_STEPS)) %>%
    mutate(category=predict(object=model, .)) %>%
    ggplot(aes(x=TIA1_L1_CLIP, y=LARP4B_exon_CLIP, fill=category)) +
    geom_tile(linejoin="bevel") +
    scale_fill_manual(values=LOC_COLORS) +
    theme_bw()

ggsave("img/lr-all-pairs-tia1l1-larp4b.pdf", width=5, height=4, dpi=300)
```

# Plot Predicted
## Ternary
```{r plot_predicted_tern, fig.width=5, fig.height=5, warning=FALSE, message=FALSE}
df_scaled %>%
    mutate(predicted=predict(model, .)) %>%
    cbind(df_tiger[, c("npco_er", "npco_cy", "npco_tg")]) %>%
    filter(category != "DF") %>%
    ggtern(aes(x=npco_er, y=npco_cy, z=npco_tg, color=predicted)) +
    geom_density_tern(size=0.5) +
    geom_point(size=0.1, alpha=0.3) +
    geom_Lline(Lintercept=0.4, linetype='dashed') + 
    geom_Rline(Rintercept=0.4, linetype='dashed') + 
    geom_Tline(Tintercept=0.4, linetype='dashed') +
    scale_color_manual(values=LOC_COLORS_DARK) +
    labs(x="ER", y="CY", z="TG", title="All Data") +
    guides(color="none") +
    theme_bw()

ggsave("img/lr-all-predict.pdf", width=5, height=5, dpi=300)
```




---

# Runtime Details
## Session Info
```{r sesh_info, echo=FALSE}
sessionInfo()
```

## Conda Environment
```{bash comment="", echo=FALSE}
if ! command -v conda &> /dev/null
then
  echo "Conda not detected."
elif [ -z "${CONDA_PREFIX}" ]
then
  echo "No active Conda environment."
else
  echo "## Conda Environment YAML"
  conda env export
fi
```
