---
title: "Initial Exploratory Data Visualizations"
author: "Mervin M Fansler"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
---

# Purpose

Here we have an initial foray into plotting some of the covariate values, such as 
RNA-binding protein associations, with respect to the readout of partition coefficients.
Since there are three partitions that are compositional, we can use ternary plots 
to represent the readout space.

# Initialization

## Libraries
```{r libs, message=FALSE, warning=FALSE}
library(magrittr)
library(tidyverse)
library(ggtern)
library(ggridges)
library(gtsummary)
library(readxl)
library(cowplot)
```

## Parameters
```{r set_params}
set.seed(20220602)

RDS_IN="data/df_tiger_clean.Rds"

LOC_COLORS=c("ER"="#4B74A6", "TG"="#E4E4CA", "CY"="#949494", "DF"="#CCCCCC")
LOC_COLORS_DARK=c("ER"="#4B74A6", "TG"="#C9C497", "CY"="#949494", "DF"="#CCCCCC")

MIN_PCON_CY=1.15
MIN_PCON_ER=1.25
MIN_PCON_TG=1.30

MIN_PCON_DIFF=0.15
```

# Data
## Loading
```{r load_data, message=FALSE}
df_tiger <- readRDS(RDS_IN)
```

# Analysis

## Sanity Checks
### Compositionality
First, validate that the partition coefficients for each gene sum to 1.

```{r pco_valid, fig.width=3, fig.height=4}
df_pcos <- df_tiger %>%
    mutate(pco_sum=pco_tg+pco_er+pco_cy) %>%
    select(gene_name, refseq_id, category, pco_sum, pco_tg, pco_er, pco_cy)

ggplot(df_pcos, aes(x=pco_sum)) +
    geom_histogram(fill='lightgrey', color='black', binwidth=0.005) +
    geom_histogram(data=filter(df_pcos, map_lgl(pco_sum, all.equal, current=1)), fill='steelblue', color='black', binwidth=0.005) +
    labs(x="Sum of Partition Coefficients", y="Genes") +
    scale_y_continuous(expand=c(0,0,0.1,0)) +
    theme_light()
```

Seems to be correct!

### Normalizaton
Normalized coefficients are derived by by dividing by the medians. Let's verify.
```{r recalc_normed_vals}
df_pcons <- df_tiger %>%
    select(gene_name, refseq_id, category, 
           pco_cy, pco_er, pco_tg, 
           npco_cy, npco_er, npco_tg) %>%
    mutate(pcon_cy=pco_cy/median(pco_cy),
           pcon_er=pco_er/median(pco_er),
           pcon_tg=pco_tg/median(pco_tg))

df_pcons %$% all.equal(npco_cy, pcon_cy)
df_pcons %$% all.equal(npco_er, pcon_er)
df_pcons %$% all.equal(npco_tg, pcon_tg)
```

Looks like there are some differences. Let's check that these really are uniformly small,
and not a few large differences.

```{r check_small_errors} 
df_pcons %$% summary(npco_cy - pcon_cy)
df_pcons %$% summary(npco_er - pcon_er)
df_pcons %$% summary(npco_tg - pcon_tg)

df_pcons %>%
    ggplot() +
    geom_histogram(aes(x=npco_cy-pcon_cy), binwidth=1e-5, position='stack',
                   fill=LOC_COLORS["CY"], color='black') +
    geom_histogram(aes(x=npco_er-pcon_er), binwidth=1e-5, position='stack',
                   fill=LOC_COLORS["ER"], color='black') +
    geom_histogram(aes(x=npco_tg-pcon_tg), binwidth=1e-5, position='stack',
                   fill=LOC_COLORS["TG"], color='black') +
    scale_y_continuous(expand=c(0,0,0.05,0)) +
    labs(x="Difference (Reported - Computed)", y="Genes") +
    theme_light()
```

They are small, but they are biased. Were the medians rounded prior to dividing?

```{r check_medians}
df_pcons %$% median(pco_cy)
df_pcons %$% median(pco_er)
df_pcons %$% median(pco_tg)
```

Yes. This is consistent with having rounded the medians to the fourth sigfig prior
to computing the normalized values. So, we should get exact matching with...

```{r rerecalc_normed_vals}
df_pcons2 <- df_tiger %>%
    select(gene_name, refseq_id, category, 
           pco_cy, pco_er, pco_tg, 
           npco_cy, npco_er, npco_tg) %>%
    mutate(pcon_cy=pco_cy/round(median(pco_cy), digits=4),
           pcon_er=pco_er/round(median(pco_er), digits=4),
           pcon_tg=pco_tg/round(median(pco_tg), digits=4))

df_pcons2 %$% all.equal(npco_cy, pcon_cy)
df_pcons2 %$% all.equal(npco_er, pcon_er)
df_pcons2 %$% all.equal(npco_tg, pcon_tg)
```

Perfect!

### Consistent Categorization 
We were told that categorization was based on the normalized partition coefficients.
We'll see how it matches up on the raw coefficients first, and then check that normalized
coefficients properly assign the categories.
```{r cat_valid_raw}
df_pcos %>%
    mutate(pco_sum=NULL) %>%
    pivot_longer(cols=starts_with("pco", ignore.case=FALSE),
                 names_to="partition", names_prefix="pco_",
                 values_to="coefficient") %>%
    group_by(gene_name, refseq_id) %>%
    slice_max(coefficient) %>%
    ungroup() %>%
    mutate(partition=factor(toupper(partition), levels=c("ER", "TG", "CY"))) %>%
    tbl_summary(by=category, include=partition, 
                label=c(category="Category", partition="Max Partition (raw)")) %>%
    modify_header(label ~ "**Categorization**")
```

Okay, that's a little noisy, but we know that the actual categories were assigned 
using the normalized coefficients. So, let's look at them...

```{r cat_valid_norm}
df_tiger %>%
    pivot_longer(cols=starts_with("npco_"),
                 names_to="partition", names_prefix="npco_",
                 values_to="coefficient") %>%
    group_by(gene_name, refseq_id) %>%
    slice_max(coefficient) %>%
    ungroup() %>%
    mutate(partition=factor(toupper(partition), levels=c("ER", "TG", "CY"))) %>%
    tbl_summary(by=category, include=partition, 
                label=c(category="Category", partition="Max Partition (norm.)")) %>%
    modify_header(label ~ "**Categorization**")
```

This shows consistency with maximum category.

## Categorization
Let's look at the categories visualized in ternary plots.

### Raw Coefficients
The first plot shows a näive 50% cutoff line. That is, any genes outside the central
triangle have 50% of their FPKMs coming from one subcellular compartment. The second
plot includes cutoff lines that are chosen to separate the classifications.

```{r plot_cat_raw, fig.width=5, fig.height=5}
df_tiger %>%
    ggtern(aes(x=pco_er, y=pco_cy, z=pco_tg, color=category)) +
    geom_point(size=0.2, alpha=0.4) +
    geom_Lline(Lintercept=0.50, linetype='dashed') + 
    geom_Rline(Rintercept=0.50, linetype='dashed') + 
    geom_Tline(Tintercept=0.50, linetype='dashed') +
    scale_color_manual(values=LOC_COLORS_DARK) +
    labs(x="ER", y="CY", z="TG") +
    guides(color="none") +
    theme_bw()

df_tiger %>%
    ggtern(aes(x=pco_er, y=pco_cy, z=pco_tg, color=category)) +
    geom_point(size=0.2, alpha=0.4) +
    geom_Lline(Lintercept=0.38, linetype='dashed') + 
    geom_Rline(Rintercept=0.42, linetype='dashed') + 
    geom_Tline(Tintercept=0.43, linetype='dashed') +
    scale_color_manual(values=LOC_COLORS_DARK) +
    labs(x="ER", y="CY", z="TG") +
    guides(color="none") +
    theme_bw()

ggsave("img/pcos_raw_ternary.pdf", width=5, height=5, dpi=300)
```

There is some non-linearity in there, but one may not notice it if not looking closely.
This is expected because we assign categories using the normalized coefficients.

### Normalized Coefficients
Three plots show 50% cutoffs, 40% cutoffs, and empirical (best fit) cutoffs.

```{r plot_cat_norm, fig.width=5, fig.height=5}
df_tiger %>%
    ggtern(aes(x=npco_er, y=npco_cy, z=npco_tg, color=category)) +
    geom_point(size=0.2, alpha=0.4) +
    geom_Lline(Lintercept=0.5, linetype='dashed') + 
    geom_Rline(Rintercept=0.5, linetype='dashed') + 
    geom_Tline(Tintercept=0.5, linetype='dashed') +
    scale_color_manual(values=LOC_COLORS_DARK) +
    labs(x="ER", y="CY", z="TG") +
    guides(color='none') +
    theme_bw()

df_tiger %>%
    ggtern(aes(x=npco_er, y=npco_cy, z=npco_tg, color=category)) +
    geom_point(size=0.2, alpha=0.4) +
    geom_Lline(Lintercept=0.4, linetype='dashed') + 
    geom_Rline(Rintercept=0.4, linetype='dashed') + 
    geom_Tline(Tintercept=0.4, linetype='dashed') +
    scale_color_manual(values=LOC_COLORS_DARK) +
    labs(x="ER", y="CY", z="TG") +
    guides(color='none') +
    theme_bw()

df_tiger %>%
    ggtern(aes(x=npco_er, y=npco_cy, z=npco_tg, color=category)) +
    geom_point(size=0.2, alpha=0.4) +
    geom_Lline(Lintercept=0.405, linetype='dashed') + 
    geom_Rline(Rintercept=0.40, linetype='dashed') + 
    geom_Tline(Tintercept=0.41, linetype='dashed') +
    scale_color_manual(values=LOC_COLORS_DARK) +
    labs(x="ER", y="CY", z="TG") +
    guides(color='none') +
    theme_bw()

ggsave("img/pcos_norm_ternary.pdf", width=5, height=5, dpi=300)
```

#### Alternative Dot Sizes
```{r plot_cat_norm_alts, fig.width=5, fig.height=5}
df_tiger %>%
    ggtern(aes(x=npco_er, y=npco_cy, z=npco_tg, color=category)) +
    geom_point(size=0.4, alpha=0.4) +
    geom_Lline(Lintercept=0.405, linetype='dashed') + 
    geom_Rline(Rintercept=0.40, linetype='dashed') + 
    geom_Tline(Tintercept=0.41, linetype='dashed') +
    scale_color_manual(values=LOC_COLORS_DARK) +
    labs(x="ER", y="CY", z="TG") +
    guides(color='none') +
    theme_bw()

ggsave("img/pcos_norm_ternary_alt1.pdf", width=5, height=5, dpi=300)

df_tiger %>%
    ggtern(aes(x=npco_er, y=npco_cy, z=npco_tg, color=category)) +
    geom_point(size=0.8, alpha=0.6) +
    geom_Lline(Lintercept=0.405, linetype='dashed') + 
    geom_Rline(Rintercept=0.40, linetype='dashed') + 
    geom_Tline(Tintercept=0.41, linetype='dashed') +
    scale_color_manual(values=LOC_COLORS_DARK) +
    labs(x="ER", y="CY", z="TG") +
    guides(color='none') +
    theme_bw()

ggsave("img/pcos_norm_ternary_alt2.pdf", width=5, height=5, dpi=300)
```

# Covariate Plots
Here we plot some of the covariates (e.g., CLIP peak counts) using the ternary plots.
We both scale the point sizes and the color based on the covariate values. This helps
to deemphasize all the those genes that have low values in favor of those with high
values. Overall, I think the individual plots are not so clear to interpret, however, 
comparing the plots of different covariates seems to make for a nice contrast.

The ridge plots are more for modelling purposes, to explore what type of transformation
might be best to apply before input the covariate into the regression model.

## TIS11B
### Ternary Plot
```{r plot_tis11_tern, fig.width=8, fig.height=6, warning=FALSE}
df_tiger %>%
    mutate(pscore_tis11b=2*(sqrt(TIS11B_CLIP) - sqrt(mean(TIS11B_CLIP, na.rm=TRUE)))) %>%
    arrange(TIS11B_CLIP) %>%
    ggtern(aes(x=npco_er, y=npco_cy, z=npco_tg, color=TIS11B_CLIP, size=TIS11B_CLIP)) +
    geom_point(alpha=0.3, pch=16) +
    geom_Lline(Lintercept=0.4, linetype='dashed') + 
    geom_Rline(Rintercept=0.4, linetype='dashed') + 
    geom_Tline(Tintercept=0.4, linetype='dashed') +
    scale_color_viridis_c(option="B", trans="sqrt") +
    scale_size_area() +
    labs(x="ER", y="CY", z="TG") +
    guides(size='none') +
    theme_light()

ggsave("img/tis11b_ternary.pdf", width=8, height=6, dpi=300)
```

This is clearly non-cytoplasmic, with leaning toward the TG side.

### Ridge Plots
```{r plot_tis11_ridge, warning=FALSE, message=FALSE}
RBP_CLIP <- "TIS11B_CLIP"

df_tiger %>%
    ggplot(aes(x=.data[[RBP_CLIP]], y=category, fill=category)) +
    geom_density_ridges(alpha=0.95) +
    scale_fill_manual(values=LOC_COLORS) +
    labs(x="Raw Counts") +
    theme_bw()

df_tiger %>%
    mutate(pscore=2*(sqrt(.[[RBP_CLIP]]) - sqrt(mean(.[[RBP_CLIP]], na.rm=TRUE)))) %>%
    ggplot(aes(x=pscore, y=category, fill=category)) +
    geom_density_ridges(alpha=0.95) +
    scale_fill_manual(values=LOC_COLORS) +
    labs(x="Variance Stabilizing Transform") +
    theme_bw()

df_tiger %>%
    mutate(zscore=scale(.[[RBP_CLIP]])) %>%
    ggplot(aes(x=zscore, y=category, fill=category)) +
    geom_density_ridges(alpha=0.95) +
    scale_fill_manual(values=LOC_COLORS) +
    labs(x="Z-Score") +
    theme_bw()
```

## TIA1_L1
### Ternary Plot
```{r plot_tia1l1_tern, fig.width=8, fig.height=6, warning=FALSE, message=FALSE}
df_tiger %>%
    mutate(pscore_tis11b=2*(sqrt(TIA1_L1_CLIP) - sqrt(mean(TIA1_L1_CLIP, na.rm=TRUE)))) %>%
    arrange(TIA1_L1_CLIP) %>%
    ggtern(aes(x=npco_er, y=npco_cy, z=npco_tg, color=TIA1_L1_CLIP, size=TIA1_L1_CLIP)) +
    geom_point(alpha=0.3, pch=16) +
    geom_Lline(Lintercept=0.4, linetype='dashed') + 
    geom_Rline(Rintercept=0.4, linetype='dashed') + 
    geom_Tline(Tintercept=0.4, linetype='dashed') +
    scale_color_viridis_c(option="B", trans="sqrt") +
    scale_size_area() +
    labs(x="ER", y="CY", z="TG") +
    guides(size='none') +
    theme_light()

ggsave("img/tial1_ternary.pdf", width=8, height=6, dpi=300)
```

### Ridge Plot
```{r plot_tial1_ridge, warning=FALSE, message=FALSE}
RBP_CLIP <- "TIA1_L1_CLIP"

df_tiger %>%
    ggplot(aes(x=.data[[RBP_CLIP]], y=category, fill=category)) +
    geom_density_ridges(alpha=0.95) +
    scale_fill_manual(values=LOC_COLORS) +
    labs(x="Raw Counts") +
    theme_bw()

df_tiger %>%
    mutate(pscore=2*(sqrt(.[[RBP_CLIP]]) - sqrt(mean(.[[RBP_CLIP]], na.rm=TRUE)))) %>%
    ggplot(aes(x=pscore, y=category, fill=category)) +
    geom_density_ridges(alpha=0.95) +
    scale_fill_manual(values=LOC_COLORS) +
    labs(x="Variance Stabilizing Transform") +
    theme_bw()

df_tiger %>%
    mutate(zscore=scale(.[[RBP_CLIP]])) %>%
    ggplot(aes(x=zscore, y=category, fill=category)) +
    geom_density_ridges(alpha=0.95) +
    scale_fill_manual(values=LOC_COLORS) +
    labs(x="Z-Score") +
    theme_bw()
```

## PUM2 CLIP
### Ternary Plot
```{r plot_pum2_tern, fig.width=8, fig.height=6, warning=FALSE, message=FALSE}
df_tiger %>%
    mutate(pscore_tis11b=2*(sqrt(PUM2_CLIP) - sqrt(mean(PUM2_CLIP, na.rm=TRUE)))) %>%
    arrange(PUM2_CLIP) %>%
    ggtern(aes(x=npco_er, y=npco_cy, z=npco_tg, color=PUM2_CLIP, size=PUM2_CLIP)) +
    geom_point(alpha=0.3, pch=16) +
    geom_Lline(Lintercept=0.4, linetype='dashed') + 
    geom_Rline(Rintercept=0.4, linetype='dashed') + 
    geom_Tline(Tintercept=0.4, linetype='dashed') +
    scale_color_viridis_c(option="B", trans="sqrt") +
    scale_size_area() +
    labs(x="ER", y="CY", z="TG") +
    guides(size='none') +
    theme_light()

ggsave("img/pum2_clip_ternary.pdf", width=8, height=6, dpi=300)
```


### Ridge Plot
```{r plot_pum2_ridge, warning=FALSE, message=FALSE}
RBP_CLIP <- "PUM2_CLIP"

df_tiger %>%
    ggplot(aes(x=.data[[RBP_CLIP]], y=category, fill=category)) +
    geom_density_ridges(alpha=0.95) +
    scale_fill_manual(values=LOC_COLORS) +
    labs(x="Raw Counts") +
    theme_bw()

df_tiger %>%
    mutate(pscore=2*(sqrt(.[[RBP_CLIP]]) - sqrt(mean(.[[RBP_CLIP]], na.rm=TRUE)))) %>%
    ggplot(aes(x=pscore, y=category, fill=category)) +
    geom_density_ridges(alpha=0.95) +
    scale_fill_manual(values=LOC_COLORS) +
    labs(x="Variance Stabilizing Transform") +
    theme_bw()

df_tiger %>%
    mutate(zscore=scale(.[[RBP_CLIP]])) %>%
    ggplot(aes(x=zscore, y=category, fill=category)) +
    geom_density_ridges(alpha=0.95) +
    scale_fill_manual(values=LOC_COLORS) +
    labs(x="Z-Score") +
    theme_bw()
```

## Pumilio num
### Ternary Plot
```{r plot_pum_num_tern, fig.width=8, fig.height=6, warning=FALSE, message=FALSE}
df_tiger %>%
    mutate(pscore=2*(sqrt(Pumilio_3UTR_num) - sqrt(mean(Pumilio_3UTR_num, na.rm=TRUE)))) %>%
    arrange(Pumilio_3UTR_num) %>%
    ggtern(aes(x=npco_er, y=npco_cy, z=npco_tg, 
               color=Pumilio_3UTR_num, 
               size=Pumilio_3UTR_num)) +
    geom_point(alpha=0.3, pch=16) +
    geom_Lline(Lintercept=0.4, linetype='dashed') + 
    geom_Rline(Rintercept=0.4, linetype='dashed') + 
    geom_Tline(Tintercept=0.4, linetype='dashed') +
    scale_color_viridis_c(option="B", trans="sqrt") +
    scale_size_area() +
    labs(x="ER", y="CY", z="TG") +
    guides(size='none') +
    theme_light()

ggsave("img/pumilio_3utr_ternary.pdf", width=8, height=6, dpi=300)
```

### Density
```{r plot_pum_density_tern, fig.width=8, fig.height=6, warning=FALSE, message=FALSE}
df_tiger %>%
    mutate(pum_density=1000*Pumilio_3UTR_num/Anno_3UTR_length) %>%
    arrange(pum_density) %>%
    ggtern(aes(x=npco_er, y=npco_cy, z=npco_tg, 
               color=pum_density, 
               size=pum_density)) +
    geom_point(alpha=0.3, pch=16) +
    geom_Lline(Lintercept=0.4, linetype='dashed') + 
    geom_Rline(Rintercept=0.4, linetype='dashed') + 
    geom_Tline(Tintercept=0.4, linetype='dashed') +
    scale_color_viridis_c(option="B", trans="sqrt") +
    scale_size_area() +
    labs(x="ER", y="CY", z="TG") +
    guides(size='none') +
    theme_light()

ggsave("img/pumilio_density_ternary.pdf", width=8, height=6, dpi=300)
```

### Ridge Plot
```{r plot_pum_num_ridge, warning=FALSE, message=FALSE}
RBP_CLIP <- "Pumilio_3UTR_num"

df_tiger %>%
    ggplot(aes(x=.data[[RBP_CLIP]], y=category, fill=category)) +
    geom_density_ridges(alpha=0.95) +
    scale_fill_manual(values=LOC_COLORS) +
    labs(x="Raw Counts") +
    theme_bw()

df_tiger %>%
    mutate(pscore=2*(sqrt(.[[RBP_CLIP]]) - sqrt(mean(.[[RBP_CLIP]], na.rm=TRUE)))) %>%
    ggplot(aes(x=pscore, y=category, fill=category)) +
    geom_density_ridges(alpha=0.95) +
    scale_fill_manual(values=LOC_COLORS) +
    labs(x="Variance Stabilizing Transform") +
    theme_bw()

df_tiger %>%
    mutate(zscore=scale(.[[RBP_CLIP]])) %>%
    ggplot(aes(x=zscore, y=category, fill=category)) +
    geom_density_ridges(alpha=0.95) +
    scale_fill_manual(values=LOC_COLORS) +
    labs(x="Z-Score") +
    theme_bw()
```

# Combination Plots
## TIS11B and TIAL1
```{r plot_tis11_tial1, fig.width=8, fig.height=6, warning=FALSE}
df_tiger %>%
    mutate(pscore_tis11b=2*(sqrt(TIS11B_CLIP) - sqrt(mean(TIS11B_CLIP, na.rm=TRUE))),
           pscore_tial1=2*(sqrt(TIA1_L1_CLIP) - sqrt(mean(TIA1_L1_CLIP, na.rm=TRUE)))) %>%
    arrange(pmax(TIS11B_CLIP,TIA1_L1_CLIP)) %>%
    ggtern(aes(x=npco_er, y=npco_cy, z=npco_tg)) +
    geom_point(aes(color=TIS11B_CLIP, size=TIS11B_CLIP+TIA1_L1_CLIP), 
               alpha=0.60, pch=16) +
    scale_color_gradient(low="#000000", high="#FF0000", trans="log1p") +
    geom_point(aes(fill=TIA1_L1_CLIP, size=TIS11B_CLIP+TIA1_L1_CLIP), 
               alpha=0.12, pch=21) +
    scale_fill_gradient(low="#000000", high="#00FF00", trans="log1p") +
    geom_Lline(Lintercept=0.4, linetype='dashed') + 
    geom_Rline(Rintercept=0.4, linetype='dashed') + 
    geom_Tline(Tintercept=0.4, linetype='dashed') +
    scale_size_area() +
    labs(x="ER", y="CY", z="TG") +
    guides(size='none') +
    theme_dark() +
    theme(panel.background=element_rect(fill="#000000"))

ggsave("img/tis11b_tial1_ternary.pdf", width=10, height=8, dpi=300)
```


## TIS11B and HuR
```{r plot_tis11_hur, fig.width=8, fig.height=6, warning=FALSE}
df_tiger %>%
    mutate(pscore_tis11b=2*(sqrt(TIS11B_CLIP) - sqrt(mean(TIS11B_CLIP, na.rm=TRUE))),
           pscore_hur=2*(sqrt(Muk_HuR_CLIP) - sqrt(mean(Muk_HuR_CLIP, na.rm=TRUE)))) %>%
    arrange(pmax(TIS11B_CLIP,Muk_HuR_CLIP)) %>%
    ggtern(aes(x=npco_er, y=npco_cy, z=npco_tg)) +
    geom_point(aes(color=TIS11B_CLIP, size=TIS11B_CLIP+Muk_HuR_CLIP), 
               alpha=0.60, pch=16) +
    scale_color_gradient(low="#000000", high="#FF0000", trans="log1p") +
    geom_point(aes(fill=Muk_HuR_CLIP, size=TIS11B_CLIP+Muk_HuR_CLIP), 
               alpha=0.15, pch=21) +
    scale_fill_gradient(low="#000000", high="#00FF00", trans="log1p") +
    geom_Lline(Lintercept=0.4, linetype='dashed') + 
    geom_Rline(Rintercept=0.4, linetype='dashed') + 
    geom_Tline(Tintercept=0.4, linetype='dashed') +
    scale_size_area() +
    labs(x="ER", y="CY", z="TG") +
    guides(size='none') +
    theme_dark() +
    theme(panel.background=element_rect(fill="#000000"))

ggsave("img/tis11b_hur_ternary.pdf", width=10, height=8, dpi=300)
```


## TIS11B and Pum1/2
```{r plot_tis11_pum12, fig.width=8, fig.height=6, warning=FALSE}
df_tiger %>%
    mutate(pscore_tis11b=2*(sqrt(TIS11B_CLIP) - sqrt(mean(TIS11B_CLIP, na.rm=TRUE))),
           pscore_pum12=2*(sqrt(PUM_RIP12) - sqrt(mean(PUM_RIP12, na.rm=TRUE)))) %>%
    arrange(pmax(TIS11B_CLIP,PUM_RIP12)) %>%
    ggtern(aes(x=npco_er, y=npco_cy, z=npco_tg)) +
    geom_point(aes(color=TIS11B_CLIP, size=TIS11B_CLIP+PUM_RIP12), 
               alpha=0.60, pch=16) +
    scale_color_gradient(low="#000000", high="#FF0000", trans="log1p") +
    geom_point(aes(fill=PUM_RIP12, size=TIS11B_CLIP+PUM_RIP12), 
               alpha=0.15, pch=21) +
    scale_fill_gradient(low="#000000", high="#00FF00", trans="identity") +
    geom_Lline(Lintercept=0.4, linetype='dashed') + 
    geom_Rline(Rintercept=0.4, linetype='dashed') + 
    geom_Tline(Tintercept=0.4, linetype='dashed') +
    scale_size_area() +
    labs(x="ER", y="CY", z="TG") +
    guides(size='none') +
    theme_dark() +
    theme(panel.background=element_rect(fill="#000000"))

ggsave("img/tis11b_pum12_ternary.pdf", width=10, height=8, dpi=300)
```

# Conclusion

Data looks very promising to predict partition coefficients from covariate values.
Ridge peaks indicate the variance stabilizing transform is very effective. We may
want to combine the VST with a z-score to create model coefficients that are directly 
comparable.

---

# Runtime Details
## Session Info
```{r sesh_info, echo=FALSE}
sessionInfo()
```

## Conda Environment
```{bash comment="", echo=FALSE}
if ! command -v conda &> /dev/null
then
  echo "Conda not detected."
elif [ -z "${CONDA_PREFIX}" ]
then
  echo "No active Conda environment."
else
  echo "## Conda Environment YAML"
  conda env export
fi
```
